---
author: "Steven Hizmi"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE)
library(tidyverse)
library(lubridate)
```

## Assignment 2

```{r read-data, echo = FALSE}
## The echo = FALSE argument to this chunk
##   means that the code will not appear in the outut
##   but the code will run so that the data is in the session

## Read Lake Monona data
## Remove rows with missing data
monona = read_csv("../../data/lake-monona-clean-2020.csv") %>%
  drop_na()
```

### Problems

### 1

The following code makes a histogram of the `days` variable in the Lake Monona data set.

```{r problem1}
ggplot(monona, aes(x=days)) +
  geom_histogram(boundary=0,binwidth=10,color="black",fill="white")
```

What is the approximate length in days during a typical winter over the past 170 years or so that Lake Monona has been at least 50% covered with ice?

### Response

100 days

### 2

Modify the code below so that:

- one of the bars is centered at 100 days
- the width of each bar is 7 days
- the fill color is your favorite color
- the x label says "Days Closed with Ice"
- the y label says "total"
- there is a title with words of your choosing

```{r problem2}
ggplot(monona, aes(x = days)) +
  geom_histogram(boundary=0,color="white",binwidth=7,fill="red") + labs(title="Lake Monona") + 
  labs(x="Days Closed with Ice", y="total") +
scale_x_continuous(breaks=seq(0,170,10))
```

### 3

Code in the next chunk makes a line plot that shows how the variable `days` changes with time (using `year1`).

```{r problem3}
ggplot(monona, aes(x = year1, y = days)) +
  geom_line() +
  geom_hline(yintercept=0)
```

What does the line of code `geom_hline(yintercept=0)` do?

### Response

plots a horizontal line at y = 0.

### 4

Modify the code in the following chunk so that:

- There are more descriptive axis labels
- There is an informative plot title
- There is a smooth trend line that has is colored red

```{r problem4}
ggplot(monona, aes(x = year1, y = days)) +
  geom_line() + geom_smooth(method = "lm", color = "red") +
  geom_hline(yintercept=0)+ labs(title = "Days covered in Ice over time") +
  labs(y="Days Closed with Ice", x="years")
```

- Describe the pattern of the trend.
- How long was Lake Monona closed with ice in a typical year near 1850?
- How long was Lake Monona closed with ice in a typical year near the present?
- About how many days less is the period of closure with ice changing per decade?
- Has the change been fairly steady, or has the change been substantially faster or slower in some time periods?

### Response

There is a negative relationship between days covered in ice and time. As time goes on, the amount of days each year that the lake is covered in ice decreases.
Just over 100 days the lake is covered with ice near the 1850s.Near 75 days for the present. About a 5 day decrease per decade in the average ice coverage per year. The change has not been steady, there are fast and slow periods of change since 1850.

